### VIVIAN DO
### ADS502 Assignment 6
### November 30th, 2022

### Data Science Using Python and R: Chapter 10 - Page 149: Questions #11, 12, 13, & 14

**For the following exercises, work with the white_wine_training and white_wine_test data sets.**
```{r}
#import data sets
wine_train <- read.csv("~/Desktop/MSADS/ADS502/white_wine_training")
wine_test <- read.csv("~/Desktop/MSADS/ADS502/white_wine_test")

#show first 5 observations of training set
head(wine_train,5)
```
```{r}
#show first 5 observations of test set
head(wine_test,5)
```
**11) Input and standardize both the training and test data sets**
```{r}
#subset predictor variables into their own matrix
X <- subset(wine_train, select = c("alcohol", "sugar"))
X_test <- subset(wine_test, select = c("alcohol", "sugar"))

#standardize predictor variables and save the output as a dataframe
Xs <- as.data.frame(scale(X))
colnames(Xs) <- c("alcohol_z", "sugar_z")

Xs_test <- as.data.frame(scale(X_test))
colnames(Xs_test) <- c("alcohol_z", "sugar_z")
```

**12) Run k-means clustering on the training data set, using two clusters**
```{r}
#run k-means clustering with 2 clusters
kmeans01 <- kmeans(Xs, centers=2)

#save cluster membership as new variable 'cluster' and convert to factors
cluster <- as.factor(kmeans01$cluster)

#separate records into two groups based on their cluster
Cluster1 <- Xs[ which(cluster == 1), ]
Cluster2 <- Xs[ which(cluster == 2), ]
```

**13) Give the mean of each variable within each cluster and use the means to identify a "Dry wines" and a "Sweet wines" cluster**
```{r}
#show summary statistics for cluster 1 of the training set
summary(Cluster1)
```
```{r}
#show summary statistics for cluster 2 of the training set
summary(Cluster2)
```

**14) Validate the clustering results by running k-means clustering on the test data set, using two clusters, and identifying a "Dry wines" and a "Sweet wines" cluster**
```{r}
#run k-means clustering with 2 clusters on the test set
kmeans02 <- kmeans(Xs_test, centers=2)

#save cluster membership as new variable 'cluster' and convert to factors
cluster_test <- as.factor(kmeans02$cluster)

#separate records into two groups based on their cluster
Cluster1_test <- Xs_test[ which(cluster == 1), ]
Cluster2_test <- Xs_test[ which(cluster == 2), ]

#show summary statistics for cluster 1 of the test set
summary(Cluster1_test)
```
```{r}
#show summary statistics for cluster 2 of the test set
summary(Cluster2_test)
```
```
